### 2017-12-04 Project/Presentation/main.tex
Slides for presentation. [WIP]

\documentclass[10pt]{beamer}
\usepackage[style=verbose,backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\g}{\nabla}
\newcommand{\xbold}{\mathbf{x}}
\newcommand{\citecomma}{\({}^{{}_{,}}\)}
%\documentclass{article}
%\usepackage{beamerarticle}
\usetheme{CambridgeUS}
\addbibresource{main.bib}
\renewcommand{\citesetup}{\tiny}

\title[Analysis of AG Methods]{An Empirical Analysis of the Benefits of Adaptive Gradient Methods}
\subtitle{Project done as a part of \\
CS6230 - Optimization Methods in Machine Learning}

\author{Vishwak Srinivasan \and Ayushi Patel}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Introduction}
\subsection{Gradient Descent}
\begin{frame}{Gradient Descent}
\begin{itemize}
\item<1->{Gradient Descent (GD) - based methods have been a staple approach used for optimization
          in both convex and non-convex problems for many years now.
          \begin{itemize}
          \item<2->{Huge success attributed to the simplicity of the method}
          \end{itemize}
         }
\item<3->{Start with a random initialization, and follow the direction of descent.
          \begin{itemize}
          \item<4->{Direction of descent given by the negative of the gradient of a particular point.}
          \end{itemize}
         }
\uncover<5->{\[\xbold := \xbold - \eta \g f(\xbold)\]}
\end{itemize}
\end{frame}

\subsection{Modifications to Gradient Descent}
\begin{frame}{Modifications to Gradient Descent}
\begin{itemize}
\item<1->{Polyak and Nesterov have proposed methods which \textit{provide an acceleration} to the updates. % add citations
          \begin{itemize}
          \item<2->{Collectively called \textbf{momentum based methods}}
          \end{itemize}
         }
\item<3->{Gradient descent is used as \textit{mini-batch gradient descent} in practice, which leads
          to \textit{variance in the calculated mini-batch gradients}. Methods like SVRG, SAG and % add citations
          SDCA have been proposed to help solve this problem.
          \begin{itemize}
          \item<4->{Collectively called \textbf{variance reduction methods}}
          \end{itemize}
         }
\item<5->{Choice of learning rate is cumbersome process in gradient based methods. Methods like Adam,
          Adagrad, Adadelta, RMSprop and so on \textit{adapt the learning rate} through the updates. % add citations
          \begin{itemize}
          \item<6->{Collectively called \textbf{adaptive gradient methods}}
          \end{itemize}
         }
\end{itemize}
\end{frame}

\section{Adaptive Gradient Methods}
\begin{frame}{Introduction to Adaptive Gradient Methods}
\uncover<1->{General form of any adaptive gradient method is: \[\xbold_{t+1} := \xbold - \eta_{t}\g_{\xbold_{t}} f(\xbold_{t})\] where \(\xbold_{t}\) is the value of \(\xbold\) at iteration \(t\)}
\(\newline\)

\uncover<2->{Key idea of any adaptive gradient method is to \textit{modify the learning rate} based on previous updates. Following slides will discuss some ``well-known'' adaptive gradient methods used in practice briefly.}
\end{frame}

\end{document}