### 2017-12-04 Project/Presentation/main.tex
Added Optimizers slides

\documentclass[10pt]{beamer}
\usepackage[style=verbose,backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\g}{\nabla}
\newcommand{\xbold}{\mathbf{x}}
\newcommand{\citecomma}{\({}^{{}_{,}}\)}
%\documentclass{article}
%\usepackage{beamerarticle}
\usetheme{CambridgeUS}
\addbibresource{main.bib}
\renewcommand{\citesetup}{\tiny}

\title[Analysis of AG Methods]{An Empirical Analysis of the Benefits of Adaptive Gradient Methods}
\subtitle{Project done as a part of \\
CS6230 - Optimization Methods in Machine Learning}

\author{Vishwak Srinivasan \and Ayushi Patel}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\section{Introduction}
\subsection{Gradient Descent}
\begin{frame}{Gradient Descent}
\begin{itemize}
\item<1->{Gradient Descent (GD) - based methods have been a staple approach used for optimization
          in both convex and non-convex problems for many years now.
          \begin{itemize}
          \item<2->{Huge success attributed to the simplicity of the method}
          \end{itemize}
         }
\item<3->{Start with a random initialization, and follow the direction of descent.
          \begin{itemize}
          \item<4->{Direction of descent given by the negative of the gradient of a particular point.}
          \end{itemize}
         }
\uncover<5->{\[\xbold := \xbold - \eta \g f(\xbold)\]}
\end{itemize}
\end{frame}

\subsection{Modifications to Gradient Descent}
\begin{frame}{Modifications to Gradient Descent}
\begin{itemize}
\item<1->{Polyak and Nesterov have proposed methods which \textit{provide an acceleration} to the updates. % add citations
          \begin{itemize}
          \item<2->{Collectively called \textbf{momentum based methods}}
          \end{itemize}
         }
\item<3->{Gradient descent is used as \textit{mini-batch gradient descent} in practice, which leads
          to \textit{variance in the calculated mini-batch gradients}. Methods like SVRG, SAG and % add citations
          SDCA have been proposed to help solve this problem.
          \begin{itemize}
          \item<4->{Collectively called \textbf{variance reduction methods}}
          \end{itemize}
         }
\item<5->{Choice of learning rate is cumbersome process in gradient based methods. Methods like Adam,
          Adagrad, Adadelta, RMSprop and so on \textit{adapt the learning rate} through the updates. % add citations
          \begin{itemize}
          \item<6->{Collectively called \textbf{adaptive gradient methods}}
          \end{itemize}
         }
\end{itemize}
\end{frame}

\section{Adaptive Gradient Methods}
\subsection{Introduction}
\begin{frame}{Introduction to Adaptive Gradient Methods}
\uncover<1->{General form of any adaptive gradient method is: \[\xbold_{t+1} := \xbold - \eta_{t}\g_{\xbold_{t}} f(\xbold_{t})\] where \(\xbold_{t}\) is the value of \(\xbold\) at iteration \(t\)}
\(\newline\)

\uncover<2->{Key idea of any adaptive gradient method is to \textit{modify the learning rate} based on previous updates. Following slides will discuss some ``well-known'' adaptive gradient methods used in practice briefly.}
\end{frame}

\subsection{Adagrad}
\begin{frame}{Adagrad} % add citations
\uncover<1->{Define: \[G_{t} \triangleq \mathrm{diag}\left(\displaystyle \sum_{i=0}^{t} \g f(\xbold_{i}) \odot \g f(\xbold_{i})\right)\]}
\uncover<2->{Adagrad updates parameters based on the rule below:
\begin{equation}
\label{adagrad}
\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{G_{t} + \epsilon}} \odot \g f(\xbold_{t})
\end{equation}
where \(\epsilon\) is a small value to prevent divide by zero errors
}
\end{frame}

\subsection{Adadelta}
\begin{frame}{Adadelta} % add citations
\uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]
\[\mathbb{E}[g]_{t} \triangleq \rho\mathbb{E}[g]_{t-1} + (1 - \rho)g_{t}\]
\[\Delta \xbold_{t} \triangleq -\eta\displaystyle\frac{\sqrt{\mathbb{E}[\Delta \xbold]_{t-1} + \epsilon}}{\sqrt{\mathbb{E}[g]_{t} + \epsilon}}\odot\g f(\xbold_{t})\]
\[\mathbb{E}[\Delta \xbold]_{t} \triangleq \rho\mathbb{E}[\Delta \xbold]_{t-1} + (1 - \rho)\left(\Delta \xbold_{t} \odot \Delta \xbold_{t}\right)\]}
\uncover<2->{Adadelta updates parameters based on the rule below:
\begin{equation}
\label{adadelta}
\xbold_{t+1} = \xbold_{t} + \Delta \xbold_{t}
\end{equation}
where \(\epsilon\) is a small value to ensure stability/relieve divide by zero errors.
}
\end{frame}

\subsection{Adam}
\begin{frame}{Adam} % add citations
\uncover<1->{Define: \[m_{t} \triangleq \beta_{1}m_{t-1} + (1 - \beta_{1})\g f(\xbold_{t}), \hspace{3mm} v_{t} \triangleq \beta_{2}v_{t-1} + (1 - \beta_{2})\left(\g (\xbold_{t}) \odot \g (\xbold_{t})\right)\]
\[\hat{m}_{t} \triangleq \displaystyle \frac{m_{t}}{1 - \beta_{1}^{t}}, \hspace{3mm} \hat{v}_{t} \triangleq \displaystyle \frac{v_{t}}{1 - \beta_{2}^{t}}\]
}
\uncover<2->{Adam updates parameters based on the rule below:
\begin{equation}
\label{adam}
\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{\hat{v}_{t}} + \epsilon} \odot \hat{m}_{t}
\end{equation}
where \(\epsilon\) is a small value to prevent divide by zero errors
}
\end{frame}

\subsection{RMSProp}
\begin{frame}{RMSProp} % add citations
\uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]
\[\mathbb{E}[g]_{t} \triangleq 0.9\mathbb{E}[g]_{t-1} + 0.1g_{t}\]
}
\uncover<2->{RMSProp updates parameters based on the rule below:
\begin{equation}
\label{rmsprop}
\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{\mathbb{E}[g]_{t} + \epsilon}} \odot \g f(\xbold_{t})
\end{equation}
where \(\epsilon\) is a small value to prevent divide by zero errors
}
\end{frame}

\end{document}