### 2018-09-30 Assignment-1/Solns-Assgn-1.tex
Port remaining solutions to LaTeX

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\newcommand{\tr}{\mathrm{tr}}
\newcommand{\deter}{\mathrm{det}}

\title{Solutions to the Assignment - 1 : CS6230 - \\
Optimization Methods in Machine Learning}
\author{Vishwak Srinivasan\\
\texttt{CS15BTECH11043}}
\date{}

\begin{document}
\maketitle

\section*{Question 1}
\begin{flushleft}
Without loss of generality, assume that \(C\) is compact and \(D\) is just closed, since we know that \(C - D\) is closed. Now we need to show that \(\exists c^{*} \in C, d^{*} \in D\) such that \(\displaystyle c^{*}, d^{*} = \text{arg} \inf_{x \in C, y \in D} d(x, y)\). This can be seen from the definition of the compactness of \(C\) and the closedness of \(D\). 
\(\newline\)

Now construct \(z = c^{*} - d^{*}\). Let us show the following: \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z\). We prove this by contradiction:
\(\newline\)

Assume that \(\displaystyle \inf_{x \in C} x^{T}z < c^{*^{T}}z\). Since \(x\) and \(c^{*} \in C\), the convex combination of \(x\) and \(c^{*}\) must \(\in C\). Now let us find out the distance between \(x_{\epsilon} = \epsilon x + (1 - \epsilon) c^{*}\) and \(d^{*}\). This is equal to:
\begin{gather}
||x_{\epsilon} - d^{*}||^{2} = ||\epsilon x + (1 - \epsilon)c^{*} - d^{*}||^{2} = ||c^{*} - d^{*} + \epsilon(x - c^{*})||^{2} \\
\implies ||x_{\epsilon} - d^{*}||^{2} = \langle c^{*} - d^{*} + \epsilon(x - c^{*}) , c^{*} - d^{*} + \epsilon(x - c^{*}) \rangle \\
\implies ||x_{\epsilon} - d^{*}||^{2} = ||c^{*} - d^{*}||^{2} + \epsilon^{2}||x - c^{*}||^{2} + 2\epsilon\langle x - c^{*}, c^{*} - d^{*}\rangle
\end{gather}

From the assumption taken, we know that \(\langle x - c^{*} , z\rangle < 0\). The last term in equation 3 will use this fact and hence for a choice of \(\epsilon\), \(||x_{\epsilon} - d^{*}||^{2} < ||c^{*} - d^{*}||^{2}\), thereby violating the fact that \(c^{*}, d^{*}\) are the closest points from \(C, D\) respectively.
\(\newline\)

We also know that: \(\langle c^{*}, z \rangle > \langle d^{*}, z \rangle\). This can be seen by sending the \(\langle d^{*}, z \rangle\) term to the LHS, and then that becomes \(||z||^{2}\). 
\(\newline\)

Just as how we proved the claim that \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z\), we will prove \(\displaystyle d^{*^{T}}z \geq \sup_{y \in D} y^{T}z\). Assume that \(d^{*^{T}}z < \sup_{y\in D} y^{T}z\). Now, consider the distance between \(y_{\epsilon} = \epsilon y + (1-\epsilon)d^{*}\) and \(c^{*}\). We know that \(y_{\epsilon} \in D\) by convexity of \(D\).

\begin{gather}
||c^{*} - y_{\epsilon}||^{2} = ||c^{*} - \epsilon y - (1 - \epsilon)d^{*}||^{2} = ||c^{*} - d^{*} + \epsilon(d^{*} - y)||^{2} \\
\implies ||c^{*} - y_{\epsilon}||^{2} = \langle c^{*} - d^{*} + \epsilon(d^{*} - y) , c^{*} - d^{*} + \epsilon(d^{*} - y) \rangle \\
\implies ||c^{*} - y_{\epsilon}||^{2} = ||c^{*} - d^{*}||^{2} + \epsilon^{2}||d^{*} - y||^{2} + 2\epsilon\langle c^{*} - d^{*}, d^{*} - y \rangle 
\end{gather}

From the assumption taken, we know that \(\langle d^{*} - y, z\rangle < 0\). Again, the last term in equation 6 will use this fact and hence for choice of \(\epsilon\), \(||c^{*} - y_{\epsilon}||^{2} < ||c^{*} - d^{*}||^{2}\), thereby violating the fact that \(c^{*}, d^{*}\) are the closest points from \(C, D\) respectively. 
\(\newline\)

Using the above facts: \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z \geq d^{*^{T}}z \geq \sup_{y \in D} y^{T}z \implies \inf_{x \in C} x^{T}z > \sup_{y \in D} y^{T}z\). Hence proved.
\end{flushleft}

\section*{Question 2}
\begin{flushleft}
Consider the functions \(f_{1}(x) = \sqrt{|x|}\) and \(f_{2}(x) = \sqrt{|x - 1|}\). Now we know that the \(\alpha\)-sublevel sets of both \(f_{1}\) and \(f_{2}\) are convex for all \(\alpha\), but the functions \(f_{1}\) and \(f_{2}\) are themselves not convex. This makes them quasiconvex functions. 
\(\newline\)

Now take the sum of these functions. Let \(f(x) = f_{1}(x) + f_{2}(x) = \sqrt{|x|} + \sqrt{|x-1|}\). Note that this function has two local minima: one at \(x = 0\) and the other at \(x = 1\). Consider \(\alpha \in [1, \sqrt{2}]\). Note that the \(\alpha\)-sublevel sets for this choice of \(\alpha\) is not convex, hence telling us that the sum of two quasi-convex functions is not necessarily convex.
\end{flushleft}

\section*{Question 3}
\begin{flushleft}
Consider an affine transformation of \(x\) given by matrices \(A \in \mathbb{R}^{k x n}\) and \(b \in \mathbb{R}^{k}\). We know that the domain of \(g(x)\) is \(\mathbb{R}^{n}\), which is convex. Note that \(Ax\) is a matrix-vector product given by:
\begin{equation}
Ax = \begin{bmatrix} a_{1}^{T}x \\ a_{2}^{T}x \\ \vdots \\ a_{k}^{T}x \end{bmatrix}
\end{equation}
Hence, \(y = Ax + b\) can be given by \(y_{i} = a_{i}^{T}x + b_{i}\quad i \in \{1, 2, \ldots, k\}\). Using a result in Boyd and Vanderberghe (Pg. 74), the canonical log-sum-exp function - \(f(x)\) is convex in \(x\).
Now the composition of a convex and a convex function is convex: therefore \(g(x) = f(y(x)) = f(Ax + b)\) is convex.
\(\newline\)

An alternative solution would be to compute the Hessian of the function. First we have to compute the gradient. Denote \(\exp(a_{i}^{T}x + b_{i}) = E_{i}\):
\begin{equation}
\frac{\partial g}{\partial x_{j}} = \frac{\displaystyle \sum_{i=1}^{n} \exp(a_{i}^{T}x + b_{i})a_{ij}}{\displaystyle \sum_{i=1}^{k} \exp(a_{i}^{T}x + b_{i})} = \frac{\displaystyle \sum_{i=1}^{n} E_{i}a_{ij}}{\displaystyle \sum_{i=1}^{n} E_{i}}
\end{equation}

For a critical point:
\begin{equation}
\label{3-grad}
\frac{\partial g}{\partial x_{q}} = \frac{\displaystyle \sum_{i=1}^{n} E_{i}a_{iq}}{\displaystyle \sum_{i=1}^{n} E_{i}} = 0
\end{equation}

Now compute the \(p;q^{th}\) element of the Hessian (ran out of symbols).
\begin{equation}
\frac{\partial^{2} g}{\partial x_{p} \partial x_{q}} = \underbrace{\frac{\displaystyle \sum_{i=1}^{k}E_{i}a_{iq}\sum_{i=1}^{k}E_{i}a_{ip}}{\displaystyle \sum_{i=1}^{k} (E_{i})^{2}}}_{= 0 \text{ due to Eqn \ref{3-grad}}} + \frac{\displaystyle \sum_{i=1}^{k}E_{i}a_{iq}a_{ip}}{\displaystyle \sum_{i=1}^{k} E_{i}}
\end{equation}

Now since the other term is only left, we have to show \(z^{T}Hz = \sum_{p,q=1}^{n}z_{p}z_{q}H_{pq} \geq 0\) to show that the function is convex.
\begin{equation}
\frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{p=1}^{n}\sum_{q=1}^{n}\sum_{i=1}^{k}z_{p}z_{q}E_{i}a_{ip}a_{iq} = \frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{i=1}^{k}E_{i}\sum_{p=1}^{n}z_{p}a_{ip}\sum_{q=1}^{n}z_{q}a_{iq} = \frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{i=1}^{k}E_{i}\left(\sum_{p=1}^{n}z_{p}a_{ip}\right)^{2} \geq 0
\end{equation}

The last inequality comes due to the fact that \(E_{i} > 0 \forall i\), and \(x^{2} \geq 0 \forall x\).
\end{flushleft}

\section*{Question 4}
\begin{flushleft}
Proving the implication: if \(f\) is convex, then inequality in the question holds:
\(\newline\)

If \(f\) is convex: then:
\begin{equation}
\label{4_1}
f(y) \geq f(x) + \nabla f(x)^{T} (y - x)
\end{equation}
Inverting \(x\) and \(y\) we get:
\begin{equation}
\label{4_2}
f(x) \geq f(y) + \nabla f(y)^{T} (x - y)
\end{equation}

Adding equations \ref{4_1} and \ref{4_2}, we get:
\begin{equation}
\left(\nabla f(x)^{T} -  \nabla f(y)^{T}\right)(y-x) \leq 0 \implies \left(\nabla f(x)^{T} - \nabla f(y)^{T}\right) ( x - y) \geq 0
\end{equation}

Proving the implication: if the condition in the question holds, then the function is convex
\(\newline\)

Now from \((\nabla f(\mathbf{x}) - \nabla f(\mathbf{y}))^{T}(x - y ) \geq 0\). This suggests that \(\nabla f\) is increasing. Consider a function: \(g(\theta) = f(\theta\mathbf{x} + (1 - \theta)\mathbf{y}) \implies g'(0) = \nabla f(\mathbf{y})^{T}(\mathbf{y} - \mathbf{x})\).

Now we know that: \(g(1) - g(0) = \displaystyle \int_{0}^{1} g'(\theta) d\theta \geq g'(0) d\theta\). This is because \(\nabla f\) is non-decreasing as shown above. \(g(1) - g(0) \geq g'(0) = \nabla f(\mathbf{x})^{T} (\mathbf{y} - \mathbf{x}) \implies g(0) \geq g(1) + \nabla f(\mathbf{y})^{T}(\mathbf{x} -\mathbf{y})\). From this: we get: \(f(\mathbf{x}) \geq f(\mathbf{y}) + \nabla f(\mathbf{y})^{T}(\mathbf{x} - \mathbf{y})\), which is the first order condition.
\end{flushleft}

\section*{Question 5}
\subsection*{Question 5.a.}
\begin{flushleft}
We adopt a similar proof technique as used in Boyd and Vanderberghe (Pg. 74).
\(\newline\)

Consider \(X(t) = Z + tV\) where \(Z \succ 0\) and \(V \in S^{n}\) (symmetric matrices). Let \(t\) be the set of all values such that \(X(t)\) is PD.

Now since \(X(t)\) is PD, \(X(t)^{-1}\) exists, and it given by:
\begin{equation}
X(t)^{-1} = [Z + tV]^{-1} = (Z[I + tZ^{-1}V])^{-1} = [I + tZ^{-1}V]^{-1}Z^{-1}
\end{equation}

A result in Matrix perturbations states that (analogous to a negative binomial expansion):
\begin{multline}
(I + tZ^{-1}V)^{-1} = I + tZ^{-1}V + t^{2}Z^{-1}VZ^{-1}V + \ldots\\\Rightarrow (I + tZ^{-1}V)^{-1}Z^{-1} = Z^{-1} + tZ^{-1}VZ^{-1} + t^{2}Z^{-1}VZ^{-1}VZ^{-1} + \ldots
\end{multline}

Let \(g(t)\) be given by:
\begin{equation}
g(t) = \tr(X(t)^{-1}) = \tr(Z^{-1}) + t\tr(Z^{-1}VZ^{-1}) + t^{2}\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots
\end{equation}

Now, compute the derivative of \(g(t)\) w.r.t. \(t\):
\begin{gather}
\frac{\partial g(t)}{\partial t} = \tr(Z^{-1}VZ^{-1}) + 2t\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots \\
\Rightarrow \frac{\partial^{2} g(t)}{\partial t^{2}} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots \\
\Rightarrow \frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1})
\end{gather}

Denote \(W = VZ^{-1}\). Hence:
\begin{equation}
\frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1}) = 2\tr(W^{T}Z^{-1}W)
\end{equation}

If \(Z\) is PD, then \(Z^{-1}\) is PD, and \(Z^{-1}V\) and \(VZ^{-1}\) are PD. Hence: \(\frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(W^{T}Z^{-1}W) \geq 0\), since the product of PD matrices is PSD. This means that \(g(0) = \tr(Z)\) is convex and \(Z \in S_{n}^{++}\).
\end{flushleft}

\subsection*{Question 5.b.}
\begin{flushleft}
Using a similar approach to 5.a, consider \(Y(t) = Z + tV\) where \(Z \succ 0\) and \(V \in S^{n}\).
\begin{multline}
Y(t) = Z + tV = Z^{0.5}[I + tZ^{-0.5}VZ^{-0.5}]Z^{0.5} \Rightarrow \deter(Y(t)) = \deter(Z^{0.5})\deter([I + tZ^{-0.5}VZ^{-0.5}])\deter(Z^{0.5})\\= \deter(Z)\deter([I + tZ^{-0.5}VZ^{-0.5}]).
\end{multline}

Consider the eigenvalues of \(Z^{-0.5}VZ^{-0.5}\) to be \(\{\lambda_{1}, \lambda_{2}, \ldots, \lambda_{n}\}\). Hence the eigenvalues of \(I + tZ^{-0.5}VZ^{-0.5}\) are \(\{1 + t\lambda_{1}, 1 + t\lambda_{2}, \ldots, 1 + t\lambda_{n}\}\).
\(\newline\)

Let \(h(t) = \deter(Z)^{\frac{1}{n}}\deter([I + tZ^{-0.5}VZ^{-0.5}])^{\frac{1}{n}}\):
\begin{equation}
h(t) = \deter(Z)^{\frac{1}{n}}\deter([I + tZ^{-0.5}VZ^{-0.5}])^{\frac{1}{n}} = \deter(Z)^{\frac{1}{n}}\left(\prod_{i=1}^{n}(1 + t\lambda_{i})\right)^{\frac{1}{n}}
\end{equation}

The second term involving the product is a geometric mean. Citing a result in Boyd and Vanderberghe about the geometric mean being concave for positive numbers, we show that \(h(t)\) is concave \(\forall t \in T\). The proof is complete by considering \(t = 0\).
\end{flushleft}

\section*{Question 7}
\begin{flushleft}
Consider the derivative of the function \(f(z) = |z|\) to be
\begin{equation}
f'(z) = \begin{cases} 1 & \text{ if} z \geq 0 \\ -1 & \text{ otherwise} \end{cases}
\end{equation}

Therefore, the Hessian can be written as:
\begin{equation}
\begin{bmatrix}
2a & \pm 1 \\ \pm 1 & 2a
\end{bmatrix}
\end{equation}

We need the eigenvalues to be \(> 0\), hence by finding the characteristic polynomial and extracting the roots we get \(a > \pm 0.5\). We take the intersection of the two sets, to finally get \(a > 0.5\).
\end{flushleft}

\section*{Question 8}
\subsection*{Question 8.a.}
\begin{flushleft}
\begin{equation}
\label{8_1_1}
f(\mathbf{x}) = \displaystyle - \sum_{i=1}^{n} \log(x_{i}) = - \log(x_{1}) - \log(x_{2}) - \ldots - \log(x_{n})
\end{equation}

Calculating the gradient of equation \ref{8_1_1} we get:
\begin{equation}
\nabla f(\mathbf{x}) = \langle -\frac{1}{x_{1}}, -\frac{1}{x_{2}}, \ldots, -\frac{1}{x_{n}} \rangle
\end{equation}

Calculating the Hessian of equation \ref{8_1_1} we get:
\begin{equation}
\displaystyle \nabla^{2} f(\mathbf{x}) [i,j] = \begin{cases}
\frac{1}{x^{2}_{i}} & \text{if } i = j \\
0 & \text{if } i \neq j
\end{cases}
\end{equation}

Now note that the eigenvalues of \(\nabla^{2} f(\mathbf{x})\) are \(\Lambda = \lbrace\frac{1}{x^{2}_{1}}, \frac{1}{x^{2}_{2}}, \ldots, \frac{1}{x^{2}_{n}}\rbrace\). Since the domain of \(f\) is \(\mathbb{R}^{n}_{++}\), we can most certainly assure that \(\nexists \lambda \in \Lambda\), such that \(\lambda \leq 0\). Hence it can be said that \(f\) is strictly convex.
\end{flushleft}

\subsection*{Question 8.b.}
\begin{flushleft}
\begin{equation}
\label{8_2_1}
f(\mathbf{x}) = \displaystyle - \sum_{i=1}^{n} x_{i} \log(x_{i}) = -x_{1}\log(x_{1}) -x_{2}\log(x_{2}) -\ldots -x_{n}\log(x_{n})
\end{equation}

Calculating the gradient of equation \ref{8_2_1} we get:
\begin{equation}
\nabla f(\mathbf{x}) = \langle -(1 + \log(x_{1})), -(1 + \log(x_{2})), \ldots, -(1 + \log(x_{n})) \rangle
\end{equation}

Calculating the Hessian of equation \ref{8_2_1} we get:
\begin{equation}
\displaystyle \nabla^{2} f(\mathbf{x}) [i,j] = \begin{cases}
-\frac{1}{x_{i}} & \text{if } i = j \\
0 & \text{if } i \neq j
\end{cases}
\end{equation}

Note that this matrix is diagonal and the eigenvalues of this matrix are \(\Lambda = \lbrace\frac{-1}{x_1}, \frac{-1}{x_2}, \ldots, \frac{-1}{x_n}\rbrace\). Since the domain of \(f\) is \(\mathbb{R}^{n}_{+}\), we can definitely tell that the eigenvalues are all less than \(0\), hence telling us that the function is non-convex.
\end{flushleft}
\section*{Question 9}
\subsection*{Question 9.a.}
\begin{flushleft}
\textit{Claim: } \(\displaystyle f(\mathbf{x}) = \frac{1}{\mathbf{x}^T P \mathbf{x} + 1}\) is quasi-concave for \(P \in \mathbb{S}^{n}_{++}\). 
\(\newline\)

\textit{Proof: } Consider the \(\alpha\)-superlevel sets of \(f(\mathbf{x})\). Denote it by \(\mathcal{U}_{\alpha}\).
\begin{equation}
\mathcal{U}_{\alpha} = \lbrace \mathbf{x} : \frac{1}{\mathbf{x}^{T} P \mathbf{x} + 1} \geq \alpha \rbrace
\end{equation}

Assuming \(\alpha > 0\), we get: \(\displaystyle \frac{1}{\alpha} - 1 \geq \mathbf{x}^{T} P \mathbf{x} \geq 0 \implies \alpha \in (0, 1)\). Consider \(\mathbf{y}\) also such that \(\displaystyle \frac{1}{\alpha} - 1 \geq \mathbf{y}^{t} P \mathbf{y}\). Now for all \(\theta \in [0,1]\), \(\mathbf{z} = \theta\mathbf{x} + (1 - \theta)\mathbf{y}\). We will show that: \(\mathbf{z}^{T} P \mathbf{z} \leq \frac{1}{\alpha} - 1\). We know that \(\displaystyle \mathbf{z}^{T} P \mathbf{z} \leq \theta^{2} \mathbf{x}^{T} P \mathbf{x} + (1 - \theta)^2 \mathbf{y}^{T} P \mathbf{y} \leq \theta^{2} \left(\frac{1}{\alpha} - 1\right) + (1 - \theta)^2 \left(\frac{1}{\alpha} - 1\right) = \frac{1}{\alpha} - 1\). Hence we now know that the \(\alpha\)-superlevel set of \(f(\mathbf{x})\) for \(\alpha \in (0, 1)\) is convex.
\(\newline\)

Now if \(\alpha = 1\) then, we get \(\mathbf{x}^{T} P \mathbf{x} \leq 0\), and since \(P \in \mathbb{S}^{n}_{++}\), \(\mathbf{x}^{T} P \mathbf{x} = 0\) is only possible, and \(\mathbf{x} = \mathbf{0} \implies \mathcal{U}_{1} = \{\mathbf{0}\}\). A singleton set is convex too.
\(\newline\)

Now if \(\alpha > 1\) then, we get \(\mathbf{x}^{T} P \mathbf{x} \leq \frac{1}{\alpha} - 1 < 0 \implies \mathbf{x}^{T} P \mathbf{x} < 0\). We know that \(\nexists \mathbf{x}\) for which this holds \(\implies U_{\alpha > 1} = \emptyset\) , and we know that an empty set is convex too.
\(\newline\)

Consider \(\alpha = 0\). Now we get \(1 \geq 0\), which is true for all \(\mathbf{x} \in \mathbb{R}^{n}\), hence \(\mathcal{U}_{0} = \mathbb{R}^{n}\), which is convex.
\(\newline\)

Last case, consider \(\alpha < 0\). Now we get \(\mathbf{x}^{T} P \mathbf{x} \geq \displaystyle \frac{1}{\alpha} - 1\). Since \(\alpha < 0 \implies \frac{1}{\alpha} < 0 \implies \frac{1}{\alpha} - 1 < 0\). But we know that \(\mathbf{x}^{T} P \mathbf{x} \geq 0\), and hence \(\mathcal{U}_{\alpha < 0} = \mathbb{R}^{n}\) which is convex.
\(\newline\)

Summarizing the cases:
\begin{center}
\begin{tabular}{| c | c |}
\hline
Choice of \(\alpha\) & \(\mathcal{U}_{\alpha}\) \\
\hline
\hline
\(\alpha > 1\) & \(\emptyset\) \\
\hline
\(\alpha = 1\) & \(\{\mathbf{0}\}\) \\
\hline
\(\alpha \in (0, 1)\) & A convex set \\
\hline
\(\alpha = 0\) & \(\mathbb{R}^{n}\) \\
\hline
\(\alpha < 0\) & \(\mathbb{R}^{n}\) \\
\hline
\end{tabular}
\end{center}

From the aforementioned cases, it should be obvious that no matter what \(\alpha\) we choose, we get a convex superlevel set, hence proving the claim that \(f(\mathbf{x})\) is quasi-concave.

\end{flushleft}
\subsection*{Question 9.b.}
\begin{flushleft}
Using the composition property stated in Pg. 102 in Boyd and Vanderberghe, where \(A = \mathbb{I}, b = \mathbf{0}\) and \(\mathbf{c} = \mathbf{a}, d = 1\) (given in this question) and the facts that:
\begin{itemize}
\item \(f(\mathbf{x}) = \mathbf{x}^{T} P \mathbf{x}\) is convex, and hence quasiconvex
\item \(\displaystyle f\left(\mathbf{\frac{x}{\mathbf{a}^{T}\mathbf{x} + 1}}\right) = \left(\frac{\mathbf{x}^{T} P \mathbf{x}}{(\mathbf{a}^{T}\mathbf{x} + 1)^2}\right)\) is hence quasiconvex from the fact mentioned above in Boyd and Vanderberghe.
\end{itemize}
\end{flushleft}

\end{document}